---
title: "eda_notebook"
author: "Jordyn Walton"
date: '2020-03-29'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(naniar)
library(caret)
library(GGally)
```

## R Markdown
This is a notebook for exploratory analysis of the training set.

```{r}
train_vital_data <- readRDS("train_vital_data.RDS")
train_label <- readRDS("train_SepsisLabel_A12H.RDS")

dim(train_vital_data)
```


`r nrow(train_vital_data)` individuals with `r ncol(train_vital_data)` variables – including demographics, vital signs and lab values

```{r}
sum(train_label$SepsisLabel)

sum(train_label$SepsisLabel)/nrow(train_vital_data)
```

1366 have sepsis (5.51%)


```{r}
demographics = names(train_vital_data)[1:6]
# > demographics
# [1] "patientid"   "Age"         "Gender"      "Unit1"       "Unit2"
# [6] "HospAdmTime"
demInc = demographics[2:3] #only including variables inherent to the patient


vitalSigns = names(train_vital_data)[7:(6+8*4)]
Lab = names(train_vital_data)[39:length(names(train_vital_data))]

train_vital_df = train_vital_data[,colnames(train_vital_data)%in% c(demInc,vitalSigns,Lab)]


train_corr = cor(train_vital_df)
train_corr[1:6,1:6]
```
Mostly NAs



```{r}
train_corr = cor(train_vital_df,use = "pairwise.complete.obs")
train_corr[1:6,1:6]
```




#how much is missing in the dataset

```{r}
head(train_vital_data)

```



```{r}
vis_miss(train_vital_data,warn_large_data = FALSE)
```

A high percentage of EtCO2 is missing




```{r}
vis_miss(train_vital_data[,grep("mean",colnames(train_vital_data))],warn_large_data = FALSE)

```


```{r}
vis_miss(train_vital_data[,grep("min",colnames(train_vital_data))],warn_large_data = FALSE)

```


```{r}
vis_miss(train_vital_data[,grep("max",colnames(train_vital_data))],warn_large_data = FALSE)

```


```{r}
vis_miss(train_vital_data[,grep("min",colnames(train_vital_data))],warn_large_data = FALSE)

```



```{r}
var_missing = miss_var_summary(train_vital_data,order=T)
var_missing[1:10,]
remove_miss_var = var_missing$variable[var_missing$pct_miss >=20]
remove_miss_var
```

If I set the threshold to remove variables with more than 90% missingness, I am able to remove `r length(remove_miss_var)` variables.
```{r}
length(remove_miss_var)
```

This leads to a dataset with 27 variables.

```{r}
train_vital_df_rem_miss = train_vital_data[,!colnames(train_vital_data) %in% remove_miss_var]
ncol(train_vital_df_rem_miss)
```

I also would like to work with complete cases in both the test and training set.

```{r}
sum(complete.cases(train_vital_df_rem_miss))
```
does this change a lot about the distribution of labels

```{r}
train_vital_cmp = train_vital_df_rem_miss[complete.cases(train_vital_df_rem_miss),]
head(train_vital_cmp)
```

```{r}
sum(train_label$SepsisLabel[train_label$patientid %in% train_vital_cmp$patientid])
```

```{r}
1078/20281
```
5.3% positive cases - similar to before


Do any variables have zero or near zero variance? No

```{r}
nzv <- nearZeroVar(train_vital_cmp, saveMetrics= TRUE)
sum(nzv$zeroVar)
sum(nzv$nzv)
```

I will only consider analyzing the data excluding patientid and hospital admit time
```{r}
train_vital_cmp_nadm = train_vital_cmp[,!colnames(train_vital_cmp) %in% c("patientid","HospAdmTime")]
```


Do any variables have high colinearity?

```{r}
train_corr = cor(train_vital_cmp_nadm)
train_corr[1:6,1:6]
```


Visualizations

```{r}
ggcorr(train_vital_cmp_nadm[,sort(colnames(train_vital_cmp_nadm))],
       method = c("pairwise.complete.obs","pearson"),label_size = 4)
```

```{r}
ggpairs(train_vital_cmp_nadm,columns = c(1,2,3,4,5,6))

```


Columns to remove

```{r}
remove_cor <- findCorrelation(train_corr,cutoff = 0.9)
length(remove_cor)
```

This involves getting rid of `r length(remove_cor)` variables, including

```{r}
remove_cor_var = colnames(train_vital_cmp_nadm)[remove_cor]
remove_cor_var
```


Now to examine the variables that remain,

```{r}
train_vital_cmp_nadm_reduced = train_vital_cmp_nadm[,-remove_cor]
colnames(train_vital_cmp_nadm_reduced)
```

Quick question: how many of the initial variables are truly removed?
 All demographics remain -ie age and gender
 

 
```{r}
sort(names(train_vital_cmp_nadm_reduced))
dim(train_vital_cmp_nadm_reduced)
```
 

 
```{r}
head(train_vital_cmp_nadm_reduced)
```


```{r}
train_vital_data_reduced = train_vital_cmp[,!colnames(train_vital_cmp) %in% remove_cor_var]
head(train_vital_data_reduced)
saveRDS(train_vital_data_reduced,"train_vital_data_reduced.RDS")
```

#after those reductions

```{r}
vis_miss(train_vital_data_reduced)
```


```{r}
train_label_reduced = train_label[train_label$patientid %in% train_vital_data_reduced$patientid,]
dim(train_label_reduced)

saveRDS(train_label_reduced,"train_SepsisLabel_A12H_red.RDS")
```



